% ══════════════════════════════════════════════════════════════
%  APPENDIX A: PLAYER INJURIES AND WORKLOAD
% ══════════════════════════════════════════════════════════════

\section{Player Injuries and Workload}
\label{app:injuries}

The welfare concern most frequently raised about the directive is
player health. Longer matches mean more minutes of high-intensity
exertion at the point when fatigue peaks. This appendix reports a
comprehensive analysis using Transfermarkt injury records across
nine leagues and 23,255~player-seasons. Table~\ref{tab:injuries}
presents the results.


\paragraph{Injury counts.}
Total injuries per player-season \textit{decreased} in treated
leagues relative to controls ($\hat{\beta} = -0.090$, SE~$= 0.032$,
$p = 0.005$, pre-treatment mean~$= 1.89$). The decline was driven
entirely by non-muscular injuries---a category dominated by contact
injuries (tackles, collisions) and ligament damage---which fell
significantly ($\hat{\beta} = -0.083$, $p = 0.005$). Muscular
injuries---the fatigue-sensitive category encompassing hamstring
strains, calf tears, and quadriceps injuries---were completely
unchanged ($\hat{\beta} = -0.007$, $p = 0.73$). The share of
injuries classified as muscular was unchanged ($\hat{\beta} = -0.001$,
$p = 0.90$). No positional heterogeneity emerged: muscular injury
rates were unchanged for attackers/midfielders ($\hat{\beta} = -0.011$,
$p = 0.67$) and defenders/goalkeepers
($\hat{\beta} \approx 0$, $p = 0.999$). The null on muscular
injuries is well-powered: the design detects a 10\% effect with
89\% power.

\paragraph{Severity.}
Total days absent was unchanged ($\hat{\beta} = 0.89$, $p = 0.73$,
pre-treatment mean~$= 66.5$ days). Games missed was unchanged
($\hat{\beta} = 0.33$, $p = 0.29$). High-severity injuries were
unchanged ($\hat{\beta} = -0.019$, $p = 0.19$). Injury recurrence
was unchanged ($\hat{\beta} = -0.017$, $p = 0.17$). Per-1,000-hour
injury rates were unchanged for both total ($\hat{\beta} = 0.15$,
$p = 0.93$) and muscular injuries ($\hat{\beta} = 0.24$, $p = 0.51$).

\paragraph{Workload resolution.}
Per-match minutes played rose by $0.76$~minutes ($p < 0.001$),
confirming individual matches are longer. But season-level aggregates
tell the opposite story: total minutes per player-season fell by
70.5~minutes ($p = 0.009$) and total matches fell by 1.01
($p < 0.001$). Average days between matches was unchanged
($p = 0.91$). Clubs responded to longer matches with greater
squad rotation---enabled by the permanent adoption of five
substitutions---keeping cumulative workload roughly constant.

\paragraph{Interpretation.}
The decline in non-muscular injuries likely reflects concurrent
trends---less physical play (documented in the fouls analysis), VAR's
deterrent effect on dangerous challenges, improvements in medical
protocols---rather than a direct effect of the directive. We report it
as context for the muscular-injury null, not as evidence that the
directive improved player safety. The policy-relevant conclusion is
that the directive's benefits (more drama, unchanged competitive
balance) came without measurable cost to player welfare.

\input{tables/tab_injuries}

