% ══════════════════════════════════════════════════════════════
%  SECTION 3: DATA
% ══════════════════════════════════════════════════════════════

\section{Data}
\label{sec:data}

Our analysis draws on three data sources. The match panel provides
league-season structure and baseline covariates. Stoppage-time data
supply the primary dependent variable. Shot-level data support the
displacement decomposition.

\subsection{Match Panel}

We construct a panel of 53,429 matches spanning 15~European leagues
and 12~seasons (2014--15 through 2025--26) from football-data.co.uk.
For each match we observe the date, final and half-time scores, total
shots and shots on target, fouls committed, yellow and red cards, and
referee identity. The panel is balanced within leagues, with no partial
seasons.

\subsection{Stoppage Time}

The primary dependent variable---second-half stoppage time in
minutes---is available for 36,413 matches (68.2\% of the panel),
scraped from ESPN's match-by-match reports. After dropping observations
with missing covariates, 36,267 matches enter the primary
difference-in-differences estimation. Coverage exceeds 94\% for the
Big~Five leagues and the Championship across all seasons; second
divisions and smaller leagues range from 40\% to 80\%.
First-half stoppage time is available for 29,317~matches and is used
for placebo tests.

Table~\ref{tab:summary_stats} reports descriptive statistics by
treatment group and period.

\input{tables/tab_summary_stats}

\subsection{Minute-Level Goals}

For the goal-timing analysis and displacement decomposition, we
supplement the match panel with shot-level data from Understat, which
records the minute, expected-goals value, and outcome of every shot
in the Big~Five leagues and the Championship. The sample comprises
388,947~shots (of which 43,116 are goals) across six leagues from
2014--15 through 2024--25. Understat does not cover other second
divisions or smaller leagues, so the displacement analysis is
restricted to these six leagues.
