% ══════════════════════════════════════════════════════════════
%  SECTION 5: IDENTIFICATION AND ROBUSTNESS
% ══════════════════════════════════════════════════════════════

\section{Identification and Robustness}
\label{sec:strategy}

The directive did not produce a clean experiment. Thirteen of fourteen
leagues shifted in the predicted direction. The binary
treatment--control distinction we impose---Big~Five top-flight leagues
as treated, all others as controls---is a pragmatic simplification of a
continuous phenomenon. This section develops three complementary
designs: a binary DiD that yields a conservative lower bound, a
within-country specification that tightens identification, and a
dose-response framework that treats the enforcement gradient itself as
the object of interest.


% ──────────────────────────────────────────────────────────────
\subsection{Binary Difference-in-Differences}
\label{subsec:binary_did}
% ──────────────────────────────────────────────────────────────

The primary specification pools all 15~leagues across 12~seasons:
%
\begin{equation}\label{eq:did}
  Y_{ilt} = \alpha + \beta \cdot (\text{Post}_t \times \text{Treated}_l)
  + \mu_l + \gamma_t + \varepsilon_{ilt}
\end{equation}
%
where $\text{Post}_t$ equals one for seasons beginning 2023--24 or
later; $\text{Treated}_l$ equals one for the Big~Five top-flight
leagues; $\mu_l$ and $\gamma_t$ are league and season fixed effects.
Because adoption was effectively simultaneous---domestic seasons began
between August 11 and August 19, 2023---this is a two-period DiD with
a common treatment date \citep{goodmanbacon2021difference,
callaway2021difference}. We report wild cluster bootstrap $p$-values
with 999~Rademacher weight replications \citep{cameron2008bootstrap}.

\input{tables/tab_first_stage}

Table~\ref{tab:first_stage} reports three specifications. The
all-leagues estimate is $\hat{\beta} = 0.734$ (SE~$= 0.278$,
$p = 0.007$; WCB $p = 0.036$). This is a \textbf{conservative lower
bound}: control leagues increased stoppage time by 1.30~minutes,
absorbing 83\% of the treated-group shift of 1.57~minutes. The DiD
captures only the differential between high-dose and medium-dose
leagues, not the full treatment effect.

The within-country specification (England, Germany, France, Italy;
$N = 23{,}728$; 8~clusters) yields $\hat{\beta} = 0.868$
($p = 0.033$; WCB $p = 0.074$), 17\% larger once cross-country
controls are removed. The Big~Five pre-post estimate
($\hat{\beta} = 1.575$, $p < 0.001$) provides a descriptive upper
bound. The truth lies in $[0.734, 1.575]$.


% ──────────────────────────────────────────────────────────────
\subsection{Dose-Response Framework}
\label{subsec:dose_response}
% ──────────────────────────────────────────────────────────────

The enforcement gradient documented in Section~\ref{sec:first_stage}
and Figure~\ref{fig:enforcement_gradient} is not a nuisance for
identification---it is the finding the binary DiD was designed to
approximate. Rather than forcing leagues into treated and control
bins, we can exploit the continuous variation in enforcement intensity:
%
\begin{equation}\label{eq:dose}
  Y_{ilt} = \alpha + \delta \cdot (\text{Post}_t \times d_l)
  + \mu_l + \gamma_t + \varepsilon_{ilt}
\end{equation}
%
where $d_l$ is league $l$'s enforcement intensity, measured as
Cohen's~$d$ for the change in second-half stoppage time
(Table~\ref{tab:enforcement}). The coefficient $\delta$ estimates the
change in outcome~$Y$ associated with a one-standard-deviation increase
in enforcement.

This specification sidesteps the contamination problem inherent in the
binary DiD: leagues with intermediate doses (the Scottish Premiership
at $d = 0.83$, the Championship at $d = 0.59$) contribute information
proportional to their actual enforcement shift rather than being
classified as untreated. The specification is estimable with
match-level data, but the dose is measured at the league level,
leaving only 14~effective observations on the dose variable.
Parametric estimation with so few dose clusters requires caution:
the functional form is imposed, and influential observations
(Serie~B at the floor, the Premier League at the ceiling) can
dominate.

We present the dose-response framework as a complement to the binary
DiD, not a replacement. The binary DiD has a well-understood
interpretation as a lower bound under control-group contamination.
The dose-response specification makes the institutional gradient
explicit and clarifies that the binary DiD captures one point on a
continuous surface. The two designs bracket the result: the binary DiD
is conservative; the dose-response specification exploits more
information but demands more from the functional form.


% ──────────────────────────────────────────────────────────────
\subsection{Event Study and Pre-Trends}
\label{subsec:event_study}
% ──────────────────────────────────────────────────────────────

Figure~\ref{fig:event_study} plots the dynamic DiD coefficients with
2022--23 as the reference period. Post-treatment coefficients are
$+0.525$ (2023--24), $+0.394$ (2024--25), and $+0.497$ (2025--26):
stable with no reversion, the signature of a permanent institutional
change.

Honest assessment requires acknowledging a consequential pre-trend
failure. A joint $F$-test yields $F = 9.007$ ($p < 0.001$). The
\citet{rambachan2023more} HonestDiD framework yields $\bar{M}^* = 0.11$:
permitting post-treatment trend deviations exceeding 0.11~minutes per
period causes the confidence interval to include zero. By the strict
criterion of \citet{roth2022pretest}, the binary DiD does not survive
trend extrapolation.

We proceed for four reasons, none requiring dismissal of this evidence.
(i)~The within-country specification absorbs country-level secular
trends and survives ($\hat{\beta} = 0.87$, $p = 0.033$).
(ii)~The mechanism is directly observable: the directive was publicly
announced, recorded in every match report, and measurable at the
individual referee level. (iii)~The pre-trend failure has identifiable
structural sources---differential VAR adoption timelines,
heterogeneous match tempo evolution---that the within-country design
removes. (iv)~The displacement result holds within treated leagues
alone, requiring no control group.


% ──────────────────────────────────────────────────────────────
\subsection{Goodman-Bacon Decomposition}
\label{subsec:bacon}
% ──────────────────────────────────────────────────────────────

Figure~\ref{fig:goodman_bacon} presents the
\citet{goodmanbacon2021difference} decomposition. In our setting,
treatment timing is simultaneous rather than staggered, so the
decomposition serves a diagnostic rather than corrective purpose: it
confirms that no single bilateral comparison dominates the TWFE
estimate. The weighted average across 45~bilateral comparisons is
0.718, close to the TWFE coefficient of 0.734. Bilateral estimates
range from $-0.735$ to $+2.220$, with the within-England pair
($+1.117$) at the top and several Serie~A bilaterals in the negative
range---reflecting its muted response relative to nominally untreated
leagues.


% ──────────────────────────────────────────────────────────────
\subsection{Robustness}
\label{subsec:robustness}
% ──────────────────────────────────────────────────────────────

Six tests probe the primary estimate's fragility.

\paragraph{Specification curve.} Across 420~specifications varying
sample, fixed effects, outcome definition, and clustering, the median
coefficient is 0.669; 88.8\% are significant at the 5\% level and
95.0\% are positive (Figure~\ref{fig:spec_curve}).

\paragraph{Permutation inference.} Random reassignment of treatment
across leagues yields a two-sided $p = 0.045$ from 1,000 permutations
(Figure~\ref{fig:permutation}).

\paragraph{Leave-one-out.} Dropping each Big~Five league, the
coefficient ranges from 0.587 (excluding the Premier League) to 0.972
(excluding Serie~A). All five remain significant at the 1\% level.

\paragraph{First-half placebo.} The DiD on first-half stoppage is
$\hat{\beta} = 0.550$ ($p < 0.001$)---positive but smaller than the
second-half estimate, consistent with enforcement attention focusing
where consequences are most visible. Half-time goals show no
treatment effect ($\hat{\beta} = -0.024$, $p = 0.30$).

\paragraph{Excluding 2022--23.} Dropping the transition season
strengthens the estimate to $\hat{\beta} = 0.810$ ($p < 0.001$).

\paragraph{Anticipatory effects.} Separating the partial-treatment
(2022--23) from the full-treatment period estimates an anticipation
share of 42.1\%, consistent with the World Cup's mid-season timing.

