% ══════════════════════════════════════════════════════════════
%  SECTION 3: GOAL DISPLACEMENT
% ══════════════════════════════════════════════════════════════

\section{Goal Displacement}
\label{sec:displacement}

The previous section documented that matches got longer. This section
asks what the additional minutes produced. The answer is the paper's
central finding: total goals per match were unchanged, but goals
\textit{moved}. They migrated from the closing minutes of regulation
into the expanded stoppage window, reshaping the dramatic arc of
matches without altering their aggregate output.


% ──────────────────────────────────────────────────────────────
\subsection{The Paradox: More Time, Same Goals}
\label{subsec:paradox}
% ──────────────────────────────────────────────────────────────

The natural expectation is that more competitive minutes produce more
goals. They did not. Total goals per match averaged 2.649 pre-directive
and 2.775 post-directive, and the DiD estimate is $\hat{\beta} = -0.033$
(SE~$= 0.038$, $p = 0.39$), a precisely estimated null. The 95\%
confidence interval rules out effects larger than approximately 1.5\% of
the pre-directive mean (Table~\ref{tab:onpitch_effects}).

The resolution lies in the minute-level data. Table~\ref{tab:goal_timing}
and Figure~\ref{fig:goal_timing} decompose per-minute scoring rates
across nine time bins for the treated leagues. Through the first 75
minutes, nothing changed: per-minute scoring rates fluctuated by less
than $\pm3$\%, variations indistinguishable from sampling noise.

Two bins break the pattern. The 76--84 minute window---the final
segment of regulation before stoppage---shows a per-minute scoring rate
decline of 8.6\%. The 90+ window shows an increase of 35.5\%. The
first-half stoppage window (45+) also increased, by 14.3\%, consistent
with the directive's application to both halves. A caveat: the 90+
increase is confirmed by a formal DiD ($p < 0.001$), but the 76--84
decline is descriptive---we report the raw rate change without a
bin-specific regression, and the magnitude could reflect sampling
variation. The displacement interpretation rests on the joint pattern
across bins, not on the 76--84 point estimate alone.

The quantity--intensity decomposition formalises the displacement. For
goals scored after the 90th minute, 87\% of the observed increase is a
pure quantity effect: more minutes multiplied by the pre-existing
per-minute scoring rate. The intensity residual---the portion
attributable to changes in how productive those minutes are---is 13\%,
statistically indistinguishable from zero. For total goals across the
full match, the quantity channel accounts for 69\% of the modest (and
insignificant) raw change (Table~\ref{tab:decomposition}). The match
exhibited approximately constant returns to time at the dose studied.

\input{tables/tab_onpitch_effects}
\input{tables/tab_goal_timing}
\input{tables/tab_decomposition}


% ──────────────────────────────────────────────────────────────
\subsection{A Model of Optimal Attacking Intensity}
\label{subsec:model}
% ──────────────────────────────────────────────────────────────

Why did goals move rather than multiply? The mechanism is strategic. We
sketch a simple framework that generates displacement as a prediction.

Consider a team trailing by one goal at minute~$t_0$ with $T$ minutes
of remaining play. The team chooses attacking intensity
$a(s) \in [0, \bar{a}]$ at each remaining minute~$s$, which determines
a per-minute scoring probability $\lambda a(s)$ and a per-minute
conceding probability $\mu a(s)$ (attacking creates defensive
exposure). The team operates under a fatigue constraint:
$\int_{t_0}^{t_0 + T} a(s)\, ds \leq E$, where $E$ is the total
energy budget for the remaining match. The team maximises the
probability of equalising minus the expected cost of conceding.

The first-order condition equalises the marginal value of attacking
intensity across minutes. When $T$ is small (old regime:
$T \approx 12$ minutes), the constraint binds tightly: the team sets
$a(s) = E/T$ at or near the corner $\bar{a}$, generating maximal
attacking desperation in every remaining minute. This is why the
76--84 window was the old regime's peak of frantic play.

When $T$ increases (new regime: $T \approx 18$), the same energy
budget spreads across more minutes. Optimal intensity at any given
minute $t$ \textit{falls}: $a^*(s) = E/T$ decreases in $T$. Total
expected goals are $\lambda \cdot \int a^*(s)\, ds = \lambda E$, which
is \textit{independent of $T$}: the product of intensity and time is
pinned by the energy constraint. Goals per minute fall, but minutes
increase proportionally. The total is unchanged.

The model generates three testable predictions.

\begin{enumerate}[nosep, leftmargin=*]
  \item \textit{Displacement}: per-minute scoring in the 76--84 window
  should fall, the 90+ window should rise, and total goals should be
  unchanged. \textbf{Confirmed}: the data match all three predictions.
  \item \textit{Proportional to dose}: leagues with larger $\Delta T$
  should exhibit larger displacement. \textbf{Consistent}: the
  enforcement gradient (Figure~\ref{fig:enforcement_gradient})
  correlates with on-pitch changes, though the small number of leagues
  limits formal testing.
  \item \textit{Competitive neutrality}: if both teams face the same
  $T$, the redistribution should not systematically favour one side.
  \textbf{Confirmed}: draw rates, home win rates, and competitive
  balance measures are all unchanged (Table~\ref{tab:onpitch_effects}).
\end{enumerate}

This framework is deliberately minimal. A richer model would endogenise
the leading team's defensive intensity, allow for substitution effects
(the five-substitution rule), and incorporate information revelation
about the exact amount of added time. We leave these extensions to
future work and note that the simple model's predictions align with
the data: constant total output, temporal redistribution, and
symmetric effects.


% ──────────────────────────────────────────────────────────────
\subsection{Late-Game Drama}
\label{subsec:drama}
% ──────────────────────────────────────────────────────────────

The displacement has experiential consequences. Goals scored after the
90th minute rose from 0.153 to 0.183 per match, a 19.6\% increase.
The share of matches in which the result changed during stoppage time
rose from 7.3\% to 10.3\%, a three-percentage-point increase
($p < 0.001$; Figure~\ref{fig:result_changes}). Late equalizers
increased from 5.3\% to 6.5\% ($+1.2$ pp); late winners rose from
7.5\% to 8.8\% ($+1.3$ pp). The broader category of game-changing
goals after the 85th minute rose from 17.9\% to 21.4\% ($+3.5$ pp).
More than one in five matches now features a decisive moment after the
85th minute; under the old regime, the figure was closer to one in six.

Scaled to the Big~Five leagues---which stage roughly 1,800 matches per
season---the directive produced approximately 55 additional matches per
season in which the result was determined after the 90th minute, a
roughly 40\% increase over the pre-directive baseline.


% ──────────────────────────────────────────────────────────────
\subsection{Competitive Neutrality and Physical Effects}
\label{subsec:neutrality}
% ──────────────────────────────────────────────────────────────

A policy that reshuffles the timing of goals could favour one type of
team or result. The data show no such asymmetry
(Table~\ref{tab:onpitch_effects}). The draw rate was unchanged
($+0.3$ pp, $p = 0.80$); the home win rate was unchanged ($-0.8$ pp,
$p = 0.49$); the away win rate was unchanged ($+0.5$ pp, $p = 0.61$).
At the league-season level, the standard deviation of final points was
unchanged ($+0.17$, $p = 0.80$; $N = 138$ league-seasons), and the
normalised Herfindahl--Hirschman index of competitive balance was
unchanged ($+0.0007$, $p = 0.14$). The neutrality follows from the
displacement mechanism: if the same goals simply arrive later, the
final distribution of results is unaffected.

The stoppage-time treatment effect is also uniform across game states:
$\hat{\beta} = 0.72$ when tied, $0.85$ when the home team leads,
$0.63$ when the away team leads ($p_{\text{interaction}} = 0.19$).
Referees added time mechanically, regardless of the scoreline.

One physical outcome warrants note. Total fouls declined secularly in
both treated and control leagues, but the decline was partially
arrested in treated leagues, producing a positive DiD of $+1.24$~fouls
per match ($p < 0.001$). Additional late-match minutes when fatigue
peaks generated marginally more physical confrontations. Yellow and red
cards showed no significant treatment effects, suggesting the
additional fouls were tactical rather than reckless. Player injury
rates---including the fatigue-sensitive category of muscular
injuries---were unaffected by the directive; we report this analysis in
Appendix~\ref{app:injuries}.


\bigskip
\noindent\textit{Summary.}\quad The directive added roughly
1.0--1.6~minutes of competitive play per match. Those minutes produced
goals at the pre-existing per-minute rate, displaced from the final
minutes of regulation into the expanded stoppage window. A simple
model of optimal attacking intensity under a fatigue constraint
rationalises this: the trailing team's energy budget, not time,
determines total output, so extending the horizon redistributes effort
without increasing it. The redistribution was competitively neutral
but experientially consequential: the share of matches featuring a
decisive stoppage-time moment rose from 7.3\% to 10.3\%, generating
roughly 55~additional result-changing finishes per Big~Five
season---approximately 40\% more than the old regime.

