\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{mathptmx}

\begin{document}
\thispagestyle{empty}

\begin{center}
{\large\textbf{$90{+}7'$: Goal Displacement and Norm Coordination\\
After FIFA's Stoppage-Time Directive}}

\bigskip

{[Author Name]}\\
\textit{[Affiliation]}\\
{[email]}

\bigskip

\textit{JEL codes: L83, J44}
\end{center}

\bigskip

\noindent
In November 2022, FIFA instructed match officials to fully compensate
for time lost during professional football matches---enforcing a rule
systematically under-applied for decades. Using match-level data from
15~European leagues and 36,000 matches spanning 12 seasons, we document
three findings.

First, the directive produced an enforcement gradient that tracked
institutional proximity to FIFA with remarkable precision. The Premier
League added 2.1~minutes per match; Serie~A, despite identical formal
treatment, added only 0.7~minutes; Serie~B showed no change.
Within-country comparisons---where top-flight and second-division
leagues share a federation but differ in connection to FIFA's referee
infrastructure---reproduce the gradient, confirming the signal
attenuated at each institutional boundary.

Second, the additional minutes produced goal \textit{displacement}
rather than goal creation. Total goals per match were unchanged
($\beta = -0.033$, $p = 0.39$), but goals migrated from the final
minutes of regulation into expanded stoppage time: per-minute scoring
in the 76--84 minute window fell 8.6\% while the 90+ window rose
35.5\%. The share of matches with a result-changing stoppage-time goal
rose from 7.3\% to 10.3\%. A model of optimal attacking intensity
under a fatigue constraint rationalizes the pattern: extending the
time horizon redistributes effort without increasing aggregate output.

Third, compliance was near-universal and permanent. Among referees with
sufficient pre- and post-directive observations, 95\% individually
increased their stoppage time, with no decay over three seasons. The
mechanism was focal-point coordination, not monitoring or
sanctions---the World Cup demonstration crystallized a new behavioral
norm that became self-reinforcing.

\end{document}
