% ============================================================
%  v3_data.tex — Section 3: Data
%  Soccer-only version: match panel, stoppage, minute-level goals
%  Numbers from results_v3_core.json + main.tex v2
% ============================================================

\section{Data}
\label{sec:data}

Our analysis combines three data sources covering match outcomes,
stoppage time, and minute-level goal records for 15~European football
leagues over up to 12 seasons (2014--15 through 2025--26).

\subsection{Match-Level Panel}

The primary dataset contains 53,429 matches from
football-data.co.uk, covering 15~leagues across 12~seasons. Each
match record includes the date, teams, full-time and half-time
scores, shots, fouls, corners, yellow and red cards, and the referee
name (where available). The 15~leagues comprise the five Big~Five
top-flight leagues (Premier League, Bundesliga, Ligue~1, Serie~A,
La~Liga), four corresponding second divisions (Championship,
2.~Bundesliga, Ligue~2, Serie~B), and six additional leagues
(Belgian Pro League, Eredivisie, Greek Super League, Primeira Liga,
Scottish Premiership, and Segunda Divisi\'{o}n).\footnote{Segunda
Divisi\'{o}n enters the sample for pre-directive seasons only; no
post-directive stoppage-time data are available from our source for
this league.}

League sizes vary substantially: the Championship plays 552~matches
per season (24~teams) compared to 306 in the Bundesliga (18~teams)
and 380 in other major leagues (20~teams). Season coverage is
uneven---the Big~Five leagues are available from 2014--15, while some
smaller leagues enter the sample later---but we verify that results
are robust to restricting to balanced panels.

\subsection{Stoppage Time}

Second-half stoppage time---the paper's primary dependent
variable---is available for 36,413~matches (68.2\% of the panel),
sourced primarily from ESPN match data. Of these, 36,267 enter the
primary DiD regression after dropping observations with missing
covariates.\footnote{Regression samples vary slightly across
analyses due to covariate availability. The first-stage DiD uses
$N = 36{,}267$; some robustness analyses use the parsimonious
specification ($N = 36{,}413$). Results are qualitatively identical
across sample definitions.} Stoppage-time data span the 2017--18
through 2025--26 seasons, the subset of the match panel for which
ESPN data are consistently available.

Coverage is asymmetric across leagues:
\begin{itemize}[leftmargin=*]
\item \textbf{High coverage (94\%+):} Premier League, Bundesliga,
  Ligue~1, Serie~A, La~Liga, Championship.
\item \textbf{Moderate coverage (40--80\%):} 2.~Bundesliga, Ligue~2,
  Serie~B, Eredivisie, Primeira Liga, Belgian Pro League, Greek Super
  League, Scottish Premiership.
\item \textbf{No post-directive data:} Segunda Divisi\'{o}n.
\end{itemize}

First-half stoppage time is available for 29,317~matches (54.9\%),
with similar coverage patterns. We use first-half stoppage as a
within-match benchmark and first-half goals as a placebo outcome
(Section~\ref{subsec:placebo}).

\textit{Missing data and treatment status.} A probit regression of
stoppage-time missingness on treatment indicators (Big~Five
membership, post-directive period, and their interaction) yields
near-zero coefficients on all terms ($p > 0.3$ for each). A Heckman
two-step selection correction applied to the main DiD specification
yields a coefficient within 5\% of the uncorrected value, with the
inverse Mills ratio statistically insignificant. Selective
missingness does not drive our results.

\subsection{Minute-Level Goals}

From Understat, we observe 388,947~shots (including 43,116~goals)
with minute-level timestamps for Big~Five leagues and the
Championship, spanning the 2014--15 through 2024--25 seasons. Each
shot record includes the expected goals (xG) value, shot type, body
part, and match situation.

This enables analysis of goal-timing profiles---specifically,
whether the directive displaced goals from regular time into
stoppage time---but with a coverage limitation: minute-level data
are available only for six of our 15~leagues. Goal-timing results
therefore apply to treated and partially-treated leagues; we cannot
construct a full DiD for minute-level outcomes.

\subsection{Summary Statistics}

Table~\ref{tab:summary_stats_v3} reports descriptive statistics for
the analysis sample, split by league group and period.

\input{tables/tab_summary_stats_v3}

Pre-directive, Big~Five leagues averaged 4.99~minutes of second-half
stoppage, compared to 4.32 in control leagues. Post-directive, both
groups increased---to 6.56 and 5.62, respectively---but the
Big~Five increase was larger. Total goals per match are similar
across groups and periods ($\approx 2.6$--$2.8$), previewing the
null total-goals result we report below.

Table~\ref{tab:balance} reports pre-treatment covariate balance.
Normalised differences \citep{imbens2009recent} are below $|0.10|$
for all outcomes except pre-treatment second-half stoppage time
(0.246), which the DiD specification absorbs via league fixed
effects.

\input{tables/tab_balance}
